---
title: "Bootstrapping_Estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bootstrapping_Estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH5773FALLdavi0682)
```

This function can be used to find point and $(1−\alpha)*100\%$ interval estimates for each of $\beta_0,\ \beta_1,\ \beta_2$. It will also create histograms for the sampling distribution of each of the estimates.

## Task 1
Add the following funtion in the package, which will output an invisible named list containing a data frame of the $\hat\beta$ iterations, the confidence intervals and point estimates along with iter, model and alpha.
```{r}
myboot<- function(df, model, iter, alpha){
  # Specify two dependent variables
  switch(model,
         model1= Y<-"Diameter",
         model2= Y<-"Density")

  # Point estimate
  mx0 <- matrix(c(1),nrow = 18,ncol = 1)
  mx <- matrix(c(mx0,df$MassFlux,df$HeatFlux),nrow=18,ncol=3)
  point <- solve(t(mx)%*%mx)%*%t(mx)%*%as.matrix(df[Y])

  # Bootstrap estimates
  hbetas <- matrix(NA, nrow = iter, ncol = 3)

  for(j in 1:iter){
    ind <- sample(1:18,18, replace = TRUE )

    # Create matrices using the data
    mx0 <- matrix(c(1),nrow = 18,ncol = 1)
    mx <- matrix(c(mx0,df[ind,]$MassFlux,df[ind,]$HeatFlux),nrow=18,ncol=3)

    # Use matrices to calculate estimates
    bhat <- solve(t(mx)%*%mx)%*%t(mx)%*%as.matrix(df[ind,Y])
    hbetas[j,] <- bhat
  }
  layout(matrix(1:3, nrow = 1,ncol = 3))

  # Display output
  hist(hbetas[,1],
       xlab = expression(widehat(beta)[0]),
       main = "Histogram of intercept estimate")

  hist(hbetas[,2],
       xlab = expression(widehat(beta)[1]),
       main = "Histogram of slope 1 estimate")

  hist(hbetas[,3],
       xlab = expression(widehat(beta)[2]),
       main = "Histogram of slope 2 estimate")

  beta0 <- hbetas[,1]
  ci1=quantile(beta0,c(alpha/2,1-alpha/2),na.rm=TRUE)

  beta1 <- hbetas[,2]
  ci2=quantile(beta1,c(alpha/2,1-alpha/2),na.rm=TRUE)

  beta2 <- hbetas[,3]
  ci3=quantile(beta2,c(alpha/2,1-alpha/2),na.rm=TRUE)

  invisible(list(hat_betas = hbetas, point_estimate  = point, "CI for Intercept"=ci1,"CI for Beta1"=ci2,
                 "CI for Beta2"=ci3, model = model, iteration_times = iter, alpha = alpha))
}
```

## Task 2 
Added to the package: MATH5773FALLdavi0682

## Task 3
For model 1, in which the dependent variable is bubble diameter. From the result of the point estimate, the multiple regression model can be written as: $y_1 = 1.0884 - 0.0002\cdot x_1 - 0.0800\cdot x_2$  

The confidence intervals of $\beta_0, \beta_1, \beta_2$ are obtained by taking the 2.5th and 97.5th percentile values from the bootstrapping estimates. 

For model 2, in which the dependent variable is bubble density. From the result of the point estimate, the multiple regression model can be written as: $y_2 = 1-1030.0326 - 57.8994\cdot x_1 + 332037.0861\cdot x_2$  

Similarly, the confidence intervals of $\beta_0, \beta_1, \beta_2$ are obtained by taking the 2.5th and 97.5th percentile values from the bootstrapping estimates.

```{r, fig.width = 6, fig.height = 4}
bubble <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/","BUBBLE2",'.xls'))
m1 <- myboot(df = bubble, model = "model1", iter = 2000, alpha = 0.05)
str(m1)

m2 <- myboot(df = bubble, model = "model2", iter = 1000, alpha = 0.05)
str(m1)
```

## Task 4 
11.30  
a. The equation of interaction model: $FORCE = \beta_0 + \beta_1 \cdot SPEED + \beta_2 \cdot RATE + \beta_3 \cdot PCTWT + \beta_4 \cdot SPEED\times RATE +  \beta_5 \cdot SPEED\times PCTWT + \beta_6 \cdot RATE\times PCTWT$

b.The slope shown below represents the change in force for every 1% increase in weight when rate is fixed at 50mm/minute and speed is fixed at 1000rpm.
$$slope_{weight(50,1000)} = \beta_3\ + 1000*\beta_5\ + 50*\beta_6$$

c.The slope shown below represents the change in force for every 1% increase in weight when rate is fixed at 150mm/minute and speed is fixed at 1000rpm.
$$slope_{weight(250,1000)} = \beta_3\ + 1000*\beta_5\ + 150*\beta_6$$
d.$FORCE = 343.1 + 0.01812 \cdot SPEED + 1.988 \cdot RATE + 9.375 \cdot PCTWT - 0.0000125 \cdot SPEED\times RATE -  0.000375 \cdot SPEED\times PCTWT + 0.0125 \cdot RATE\times PCTWT$
```{r}
drillmetal <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/","DRILLMETAL",'.xls'))
names(drillmetal)
lm.drill <- lm(FORCE ~ SPEED + RATE + PCTWT + SPEED:RATE + SPEED:PCTWT + RATE:PCTWT, data = drillmetal)
summary(lm.drill)
```

e. $\beta_6$ needs to be tested to determine if the slopes in parts b and c are different. The $p-value$ is 0.232036, which is greater than $\alpha$ = 0.05, suggesting that we failed to reject the null hypothesis that the slops are indifferent in part b and c. Therefore, the slops in b and c are not statistically different. 


## Task 5
11.11  
a. The test statistic and p-value of the test are $F$ = 28.65 and $p-value$ = 0. Since $p-value$ is smaller than $\alpha$ = 0.05, there is sufficient evidence to conclude that the model fit is statistically useful.
```{r}
boilers <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/","BOILERS",'.xls'))
names(boilers)
lm.boilers <- lm(`Man-HRs` ~ Capacity + Pressure + Capacity:Pressure, data = boilers)
summary(lm.boilers)
```

b. The test statistic and two-tailed $p-value$ are $t$ = 2.233 and $p-value$ = 0.0327. The upper-tailed $p-value$  = 0.032662/2 = 0.0164, which is smaller than $\alpha$ = 0.05. Therefore, we can reject the null hypothesis and conclude that the rate of change of man-hours with capacity increases as the design pressure increases.  

c. The slope shown below represents the change in man-hours for every 1-psi increase in design pressure when boiler capacity is 750 thousand pounds/hour.
$$slope_{pressure(750)} = \beta_2\ + 750*\beta_3 = -1.53 + 0.003*2*750 = 0.72$$

## Task 6
11.12  
a. the scatter plot shows that usage appears to have a curvilinear relation with the size of the home, which provides some evidence for including the quadratic term in the model.
```{r, fig.width = 6, fig.height = 4}
kwhrs <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/","KWHRS",'.xls'))
plot(kwhrs$SIZE,kwhrs$USAGE, main="Scatterplot", xlab="SIZE", ylab="USAGE", pch=19)
```

b. $Usage = -806.7 + 1.962*Size - 0.0003*Size^2 $
```{r}
qua.kwhrs <- lm(USAGE ~ SIZE + SIZESQ, data = kwhrs)
summary(qua.kwhrs)
```

c. The graph shows that the model is a good fit to the data. Also, the $R^2_a$ = 0.9735 provides a numerical measure of it, which implies that about 97.35% of the sample variation in electrical usage
can be explained by the quadratic model after adjusting for sample size and degrees of freedom.
```{r, fig.width = 6, fig.height = 4}
size = seq(1000, 4000, 1)
plot(kwhrs$SIZE,kwhrs$USAGE, main="Scatterplot", xlab="Size ", ylab="Electrical Usage", pch=19)
lines(size, predict(qua.kwhrs,list(SIZE=size, SIZESQ=size^2)), col = "darkgreen", lwd = 3)  #Add fitted line
```

d. When $Size = 0$, the estimated electrical usage from the model is negative, thus the interpretation of the intercept $\beta_0$ is not meaningful. The estimated coefficient of $Size$, $\beta_1$ does not have a meaningful interpretation in the quadratic by itself. The sign of the coefficient of the quadratic term, $\beta_2$, implies downward concavity. The plot in part c suggests that the positive relation between electrical usage and size is diminishing, and the estimated electrical usage is leveling off as the home sizes increase beyond 2,500 square feet model.  

e. The test statistic and two-tailed $p-value$ are $F$ = 258.1 and $p-value$ = 0. The $p-value$ is smaller than $\alpha$ = 0.01. Therefore, we can reject the null hypothesis and conclude that the
overall model is a useful predictor of electrical usage.

f. The test statistic and two-tailed $p-value$ are $t$ = -10.599 and $p-value$ = 0.00000019. The upper-tailed $p-value$  = 0.00000019/2 = 0.00000009, which is smaller than $\alpha$ = 0.01. Therefore, we can reject the null hypothesis, and these is strong evidence that electrical usage increases more slowly per square foot for large homes than for small homes.

