---
title: "Project 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH5773FALLdavi0682)
```


This project includes three functions:
1. One-way table analysis
2. Two-way table analysis
3. Fisher's test

# One-way table analysis
With an input of a file name, this function will output a table, bar plots, the results for hypothesis test (all cell proportions are not significantly different), and confidence intervals for all cell values.\  

The approximate confidence intervals for the frequencies of Agree Strong and Agree Some opinions do not contain 0.25. Also, the Chi-square test results shows the X-squared of 14.8049 and the p-value = 0.001991. We can reject the null hypothesis that the opinions are evenly divided. Therefore, using alpha = 0.05, we can make the the conclusion that the opinions of Internet users are not evenly divided among the four categories.
```{r}
my_one_way <- function (filename, alpha = 0.05){
  # Import the data
  scan <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/",filename,'.xls'))

  # Form the table
  Number <- scan$NUMBER
  Category <- scan$CATEGORY
  margin_tab <- addmargins(xtabs(Number ~ CATEGORY, data = scan))

  # Bar plot
  gg <- ggplot2::ggplot(data = scan, mapping = ggplot2::aes(x = Category, y = Number, fill = Category)) +
    ggplot2::geom_bar(stat="identity", position = "dodge") +
    ggplot2::geom_text(ggplot2::aes(label = Number), vjust = -0.2, size = 5, position = ggplot2::position_dodge(0.9)) +
    ggplot2::ylim(0, max(Number)*1.1) +
    ggplot2::labs(title = "Equifax/Harris Consumer Privacy Survey", subtitle = "Scan",
                  x = "Response Category", y = "Number of Users")

  # Hypothesis test
  tab <- table(rep(Category, Number))
  freq1way <- s20x::freq1way(tab,hypothprob = rep(1/length(Category),length(Category)), conf.level = alpha)

  df <- as.data.frame(rep(Category, Number))
  names(df) <- "Response_Category"

  # Z statistics
  z = qnorm(1-alpha/2)
  phat = Number/sum(Number)

  # Confidence interval
  cil = phat-z*sqrt(phat*(1-phat)/sum(Number))
  ciu = phat+z*sqrt(phat*(1-phat)/sum(Number))
  ci = list(cil,ciu)

  conf_int = paste0("The approximate confidence interval for the frequency of ", Category, " is: (",
                    round(as.numeric(unlist(ci[1])),3), ", ",round(as.numeric(unlist(ci[2])),3), ").")

  # Output list
  list(table = margin_tab, plot = gg, hypothesis = freq1way, confidence_interval = conf_int)
}
```

```{r}
my_one_way("SCAN")
```

# Two-way analysis
With an input of a file name, this function will output a contingency table, bar plots, expected frequencies for all cells, Chi-squared test results, and a plot of chi-squared test rejection region.\ 
 
The null hypothesis is that injury rate for collision claims does not depend on Accord model. The alternative hypothesis is that injury rate for collision claims depends on Accord model. The chi-sq statistic is 1346.1, this number is significantly larger than 3.8415. Also, the p value is significantly less than 0.05. Therefore, we would reject the null hypothesis and conclude that and injury rate for collision claims depends on Accord model.
```{r}
my_two_way = function (filename, alpha = 0.05){

  # Import the data
  hybrid <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/",filename,'.xls'))

  Number <- hybrid$Number
  Model <- hybrid$Model
  Claim <- hybrid$Claim

  # Form the contingency table
  tab <- xtabs(Number ~ Model + Claim, data = hybrid)
  margin_tab <- addmargins(tab)

  nr = nrow(tab)
  nc = ncol(tab)
  n = sum(tab)

  margin_row = apply(tab,1,sum)
  margin_col = apply(tab,2,sum)

  # Bar plot
  gg1 <- ggplot2::ggplot(hybrid, ggplot2::aes(x = factor(Model), y = Number, fill = Claim, colour = Claim)) +
    ggplot2::geom_bar(stat = "identity", position = "dodge")

  gg2 <- ggplot2::ggplot(hybrid, ggplot2::aes(x = factor(Model), y = Number, fill = Claim, colour = Claim)) +
    ggplot2::geom_bar(stat = "identity")

  # Expected frequency
  expect_freq = matrix(rep(margin_col,nr), nrow=nr, ncol=nc,byrow = TRUE)*margin_row/n

  # Individual chi squared
  ind_chiq = matrix((tab-expect_freq)^2/expect_freq, nrow=nr, ncol=nc)

  # Sum of chi square
  chsqsum = sum(ind_chiq)
  chsq = chisq.test(tab)

  # Find the rejection region
  rej = qchisq(1-alpha,chsq$parameter)

  # Plot the rejection area

  # Create density curve
  x <- NULL
  curve(dchisq(x, df = chsq$parameter), from = 0, to = 8, main = 'Chi-Square Distribution', ylab = 'Density',lwd = 2)

  # Create vector of x values
  x_vector <- seq(qchisq(1-alpha,chsq$parameter), 8)

  # Create vector of chi-square density values
  p_vector <- dchisq(x_vector, df = chsq$parameter)

  # Fill in portion of the density plot from 0 to 8
  polygon(c(x_vector, rev(x_vector)), c(p_vector, rep(0, length(p_vector))), col = adjustcolor('red', alpha.f = 0.3), border = NA)

  # Output the list
  list(tab = tab, bar_plot_1 = gg1, bar_plot_2 = gg2, exp_freq = expect_freq, chi_sq = round(ind_chiq, digits = 4), chi_sq_sum = chsqsum, p_value = chsq$p.value, degree_of_freedom = chsq$parameter, reject_region = rej)

}
```

```{r}
my_two_way("HYBRID")
```

# Fisher's test
With an input of file name, this function will output a contingency table, bar plots, mosaic plot, Chi-squared test, and fisher's test.\

We use fisher’s exact test because the expected value is not greater than or equal to 5, which is the assumption of chi-square test. The Chi-square approximation could be incorrect.

The fisher’s test shows the p-value of 0.2616, which is larger than alpha = 0.05. Therefore, we fail to reject the null hypothesis that the distribution of ARI scores are not different between the two types of bonding adhesives.
```{r}
my_fishser = function (filename){

  # Import the data
  bonding <- readxl::read_excel(paste0("C:/Users/sergi/OneDrive/Desktop/MATH5773/Excel/",filename,'.xls'))

  NUMBER <- bonding$NUMBER
  ADHESIVE <- bonding$ADHESIVE
  ARIScore <- bonding$ARIScore

  # Form the contingency table
  tab <- xtabs(NUMBER ~ ADHESIVE + ARIScore, data = bonding)
  margin_tab <- addmargins(tab)

  # Bar plot
  gg1 <- ggplot2::ggplot(bonding, ggplot2::aes(x = factor(ARIScore), y = NUMBER, fill = ADHESIVE, colour = ADHESIVE)) +
    ggplot2::geom_bar(stat = "identity", position = "dodge")

  gg2 <- ggplot2::ggplot(bonding, ggplot2::aes(x = factor(ARIScore), y = NUMBER, fill = ADHESIVE, colour = ADHESIVE)) +
    ggplot2::geom_bar(stat = "identity")

  # Mosaic plot
  mos <- mosaicplot(tab, main = "Mosaic plot", color = TRUE)

  # Chi-squared test
  chi_test <- chisq.test(tab)

  # Fisher's test
  fisher_test <- fisher.test(tab)


  #Output the list
  list(tab = margin_tab, bar_plot_1 = gg1, bar_plot_2 = gg2, mpsaic_plot = mos, exp_freq = chi_test$expected, fishers_test = fisher_test, p_value = fisher_test$p.value)

}

```

```{r}
my_fishser("BONDING")
```

